---
title: "Untitled"
author: "TSA Pro"
date: "2025-03-18"
output:
  pdf_document: default
  PDF: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



```{r cars}
library(mgcv)
library(ggplot2)
library(ggfortify)
library(TSA)
library(dynlm)
library(stats)
library(PerformanceAnalytics)
library(fable)
library(reshape2)
library(astsa)
library(forecast)
library(tidyverse)

```


```{r pressure, echo=FALSE}
data = read.csv("VehicleData-1.csv", header = T)
data$DATE = as.Date(data$DATE, format =  "%m/%d/%Y")

data %>% mutate_all(~replace(., is.nan(.), 0))

head(data)

sales = ts(data$Total.Sales, start = c(2000,1,1), frequency = 12 )
orders = ts(data$New.Orders, start = c(2000,1,1), frequency = 12)

# length(sales)
```

```{r}
sales.plot = autoplot(sales) + xlab("date") + ylab("sales") + ggtitle("Total Sales")
sales.plot

orders.plot = autoplot(orders) + xlab("data") + ylab("orders") + ggtitle("New Orders")
orders.plot

newplots = cbind(sales, orders)
new.plots = autoplot(newplots, facets = F) + scale_y_continuous(trans = "log10")
new.plots

```

```{r}
acf(sales, lag.max = 200)
pacf(sales, lag.max = 200)

acf(orders, lag.max = 200)
pacf(orders, lag.max = 200)


##### (not sure if code below is correct, to check a relation between two univariate time series variables)

## checking relation between new orders and total sales within 12 month.
total.sales  = as.numeric(sales)
new.orders = as.numeric(orders)
astsa::lag2.plot(new.orders, total.sales, 12)

## checking correlation coefficient between two time series variables: new orders and total sales
ccf(new.orders, total.sales)
```

(1) both of acf, most lags extend beyond the confidence band, indicating a significant autocorrelation at those lags, those lags are gradually declining, suggesting a long term dependency. 

(2) A seasonality also exists in both acf plots. 

(3) from the ccf, all lags are outside of confidence band, suggesting the correlation coefficient is very significant, indicating every month, there is a strong relation between new orders and total sales of the month. 

```{r}
## X-axis points converted to 0-1 scale
sales.time.pts = c(1:length(sales))
sales.time.pts = c(sales.time.pts - min(sales.time.pts))/max(sales.time.pts)

orders.time.pts = c(1:length(orders))
orders.time.pts = c(orders.time.pts - min(orders.time.pts))/max(orders.time.pts)

## Splines Trend Estimation on sales
sales.spline = gam(sales.time.pts~s(sales.time.pts))
sales.spline.fit = ts(fitted(sales.spline),start=2000,frequency=12)

## Splines Trend Estimation on orders
orders.spline = gam(orders.time.pts~s(orders.time.pts))
orders.spline.fit = ts(fitted(orders.spline), start = 2000, frequency = 12)


## plot trends, but doesnt show up in the plot, need to fix this . 

## add trend to total sales 
plot(sales,ylab="Total Sales")
lines(sales.spline.fit,lwd=2,col="brown")
abline(sales.spline.fit[1],0,lwd=2,col="blue")

## add trend to new orders
plot(orders,ylab="New Orders")
lines(orders.spline.fit,lwd=2,col="brown")
abline(orders.spline.fit[1],0,lwd=2,col="blue")

```

(1) External factors caused a dramatic drop in sales and orders in year 2009 and 2021. 

(2) it is a non linear trend in both plots, in general the trend in new orders plot is increasing, but in total sales plot, it is slightly decreasing. 

```{r}
## check seasonlity on sales
sales.model1=lm(sales~harmonic(sales))
summary(sales.model1)

sales.model2=lm(sales~harmonic(sales,2))
summary(sales.model2)

## check seasonlity on sales
orders.model1=lm(orders~harmonic(orders))
summary(orders.model1)

orders.model2=lm(orders~harmonic(orders,2))
summary(orders.model2)

```

p values of seasonality estimation of total sales and new orders are bigger than 0.05, a non significant statistics, 

```{r}
## Fit a non-parametric model for trend and linear model for seasonality on total sales
sales.har = harmonic(sales,2)
sales.gam.fit = gam(sales~s(sales.time.pts)+sales.har)
sales.dif.fit.gam = ts((sales-fitted(sales.gam.fit)),start=2000,frequency=12)
ts.plot(sales.dif.fit.gam,ylab="residual", main="Residual Process on Total sales")

## Fit a non-parametric model for trend and linear model for seasonality on new orders
orders.har = harmonic(orders,2)
orders.gam.fit = gam(orders~s(orders.time.pts)+orders.har)
orders.dif.fit.gam = ts((orders-fitted(orders.gam.fit)),start=2000,frequency=12)
ts.plot(orders.dif.fit.gam,ylab="residual", main="Residual Process on New orders")

```

```{r}
##  fitting arima model on sales 

## differencing with order 1
new.data.sales = sales

# 60 is the 20% of length of sales
sales.train = new.data.sales[1:(length(new.data.sales) - 60)]
sales.test = new.data.sales[(length(new.data.sales)- 60 + 1) : length(new.data.sales)]
# reference: lecture notes from module 2
n = length(sales.train)
norder = 7
p = c(1:norder)-1
q = c(1:norder)-1

aic = matrix(0, norder, norder)

# differencing with order 1
for (i in 1:norder) {
  for (j in 1:norder) {
    modij = arima(sales.train, order=c(p[i], 1, q[j]), method = "ML", include.mean = FALSE)
    aic[i, j] = modij$aic + 2 * (p[i] + q[j] +1)  * (p[i] + q[j]) / (n - p[i] - q[j] - 1)
  } 
}

indexp = rep(c(1:norder),norder)
indexq = rep(c(1:norder),each=norder)
indexaic = which(aic == min(aic))
porder = indexp[indexaic]-1
qorder = indexq[indexaic]-1
porder 
qorder

sales.final.diff.one = arima(sales.train, order = c(porder, 1, qorder), method = "ML")
sales.final.diff.one

```


```{r}
##  fitting arima model on orders 

## differencing with order 1
new.data.orders = orders

# 60 is the 20% of length of sales
orders.train = new.data.orders[1:(length(new.data.orders) - 60)]
orders.test = new.data.orders[(length(new.data.orders)- 60 + 1) : length(new.data.orders)]
# reference: lecture notes from module 2
n = length(orders.train)
norder = 7
p = c(1:norder)-1
q = c(1:norder)-1

aic = matrix(0, norder, norder)

# differencing with order 1
for (i in 1:norder) {
  for (j in 1:norder) {
    modij = arima(orders.train, order=c(p[i], 1, q[j]), method = "ML", include.mean = FALSE)
    aic[i, j] = modij$aic + 2 * (p[i] + q[j] +1)  * (p[i] + q[j]) / (n - p[i] - q[j] - 1)
  } 
}

indexp = rep(c(1:norder),norder)
indexq = rep(c(1:norder),each=norder)
indexaic = which(aic == min(aic))
porder = indexp[indexaic]-1
qorder = indexq[indexaic]-1
porder 
qorder

orders.final.diff.one = arima(orders.train, order = c(porder, 1, qorder), method = "ML")
orders.final.diff.one
```

```{r}
## arima evaluation on sales
sales.resid = resid(sales.final.diff.one)
plot(sales.resid, main = "Residual plot of total sales ( p =  3 & q = 0)")
acf(sales.resid, lag.max = 200,  main = "Residual ACF of total sales ( p =  3 & q = 0)")
pacf(sales.resid, lag.max = 200, main = "Residual PACF of total sales ( p =  3 & q = 0)")
Box.test(sales.resid, lag = ( 0+3+1 ), type = "Ljung-Box", fitdf = (0+3))


## arima evaluation on orders
orders.resid = resid(sales.final.diff.one)
plot(orders.resid, main = "Residual plot of New orders ( p =  5 & q = 6)")
acf(orders.resid, lag.max = 200,  main = "Residual ACF of New orders ( p =  5 & q = 6)")
pacf(orders.resid, lag.max = 200, main = "Residual PACF of New orders ( p =  5 & q = 6)")
Box.test(orders.resid, lag = ( 5+6+1 ), type = "Ljung-Box", fitdf = (5+6))

```

P value of total sales is bigger than 0.05, so no correlation. but p value of new orders is smaller than 0.05, suggesting a correlation. 





