---
title: "Untitled"
author: "TSA Pro-2"
date: "2025-04-15"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(mgcv)
library(ggplot2)
library(ggfortify)
library(TSA)
library(dynlm)
library(stats)
library(PerformanceAnalytics)
library(fable)
library(reshape2)
library(astsa)
library(forecast)
library(tidyverse)
library(rugarch)
```

```{r}
data = read.csv("VehicleData-1.csv", header = T)
data$DATE = as.Date(data$DATE, format =  "%m/%d/%Y")

data %>% mutate_all(~replace(., is.nan(.), 0))

head(data)

sales = ts(data$Total.Sales, start = c(2000,1,1), frequency = 12 )
orders = ts(data$New.Orders, start = c(2000,1,1), frequency = 12)

```


```{r}
sales.plot = autoplot(sales) + xlab("date") + ylab("sales") + ggtitle("Total Sales")
sales.plot

orders.plot = autoplot(orders) + xlab("data") + ylab("orders") + ggtitle("New Orders")
orders.plot

newplots = cbind(sales, orders)
new.plots = autoplot(newplots, facets = F) + scale_y_continuous(trans = "log10")
new.plots
```

```{r}
acf(sales, lag.max = 200)
pacf(sales, lag.max = 200)

acf(orders, lag.max = 100)
pacf(orders, lag.max = 100)


##### (not sure if code below is correct, to check a relation between two univariate time series variables)

## checking relation between new orders and total sales within 12 month.
total.sales  = as.numeric(sales)
new.orders = as.numeric(orders)
#astsa::lag2.plot(new.orders, total.sales, 12)

## checking correlation coefficient between two time series variables: new orders and total sales
#ccf(new.orders, total.sales)
```

```{r}
## X-axis points converted to 0-1 scale
sales.time.pts = c(1:length(sales))
sales.time.pts = c(sales.time.pts - min(sales.time.pts))/max(sales.time.pts)

orders.time.pts = c(1:length(orders))
orders.time.pts = c(orders.time.pts - min(orders.time.pts))/max(orders.time.pts)

## Splines Trend Estimation on sales
sales.spline = gam(sales.time.pts~s(sales.time.pts))
sales.spline.fit = ts(fitted(sales.spline),start=2000,frequency=12)

## Splines Trend Estimation on orders
orders.spline = gam(orders.time.pts~s(orders.time.pts))
orders.spline.fit = ts(fitted(orders.spline), start = 2000, frequency = 12)


## plot trends, but doesnt show up in the plot, need to fix this . 

## add trend to total sales 
plot(sales,ylab="Total Sales")
lines(sales.spline.fit,lwd=2,col="brown")
abline(sales.spline.fit[1],0,lwd=2,col="blue")

## add trend to new orders
plot(orders,ylab="New Orders")
lines(orders.spline.fit,lwd=2,col="brown")
abline(orders.spline.fit[1],0,lwd=2,col="blue")
```

```{r}
## check seasonlity on sales
sales.model1=lm(sales~harmonic(sales))
summary(sales.model1)

sales.model2=lm(sales~harmonic(sales,2))
summary(sales.model2)

## check seasonlity on sales
orders.model1=lm(orders~harmonic(orders))
summary(orders.model1)

orders.model2=lm(orders~harmonic(orders,2))
summary(orders.model2)

```

```{r}
## Fit a non-parametric model for trend and linear model for seasonality on total sales
sales.har = harmonic(sales,2)
sales.gam.fit = gam(sales~s(sales.time.pts)+sales.har)
sales.dif.fit.gam = ts((sales-fitted(sales.gam.fit)),start=2000,frequency=12)
ts.plot(sales.dif.fit.gam,ylab="residual", main="Residual Process on Total sales")

## Fit a non-parametric model for trend and linear model for seasonality on new orders
orders.har = harmonic(orders,2)
orders.gam.fit = gam(orders~s(orders.time.pts)+orders.har)
orders.dif.fit.gam = ts((orders-fitted(orders.gam.fit)),start=2000,frequency=12)
ts.plot(orders.dif.fit.gam,ylab="residual", main="Residual Process on New orders")
```

```{r}
##  fitting arima model on sales 

## differencing with order 1
new.data.sales = sales

# 60 is the 20% of length of sales
sales.train = new.data.sales[1:(length(new.data.sales) - 60)]
sales.test = new.data.sales[(length(new.data.sales)- 60 + 1) : length(new.data.sales)]
# reference: lecture notes from module 2
n = length(sales.train)
norder = 7
p = c(1:norder)-1
q = c(1:norder)-1

aic = matrix(0, norder, norder)

# differencing with order 1
for (i in 1:norder) {
  for (j in 1:norder) {
    modij = arima(sales.train, order=c(p[i], 1, q[j]), method = "ML", include.mean = FALSE)
    aic[i, j] = modij$aic + 2 * (p[i] + q[j] +1)  * (p[i] + q[j]) / (n - p[i] - q[j] - 1)
  } 
}

indexp = rep(c(1:norder),norder)
indexq = rep(c(1:norder),each=norder)
indexaic = which(aic == min(aic))
porder = indexp[indexaic]-1
qorder = indexq[indexaic]-1
porder 
qorder

sales.final.diff.one = arima(sales.train, order = c(porder, 1, qorder), method = "ML")
sales.final.diff.one

```

```{r}
##  fitting arima model on orders 

## differencing with order 1
new.data.orders = orders

# 60 is the 20% of length of sales
orders.train = new.data.orders[1:(length(new.data.orders) - 60)]
orders.test = new.data.orders[(length(new.data.orders)- 60 + 1) : length(new.data.orders)]
# reference: lecture notes from module 2
n = length(orders.train)
norder = 7
p = c(1:norder)-1
q = c(1:norder)-1

aic = matrix(0, norder, norder)

# differencing with order 1
for (i in 1:norder) {
  for (j in 1:norder) {
    modij = arima(orders.train, order=c(p[i], 1, q[j]), method = "ML", include.mean = FALSE)
    aic[i, j] = modij$aic + 2 * (p[i] + q[j] +1)  * (p[i] + q[j]) / (n - p[i] - q[j] - 1)
  } 
}

indexp = rep(c(1:norder),norder)
indexq = rep(c(1:norder),each=norder)
indexaic = which(aic == min(aic))
porder = indexp[indexaic]-1
qorder = indexq[indexaic]-1
porder 
qorder

orders.final.diff.one = arima(orders.train, order = c(porder, 1, qorder), method = "ML")
orders.final.diff.one
```

```{r}
## arima evaluation on sales
sales.resid = resid(sales.final.diff.one)
plot(sales.resid, main = "Residual plot of total sales ( p =  3 & q = 0)")
acf(sales.resid, lag.max = 200,  main = "Residual ACF of total sales ( p =  3 & q = 0)")
pacf(sales.resid, lag.max = 200, main = "Residual PACF of total sales ( p =  3 & q = 0)")
Box.test(sales.resid, lag = ( 0+3+1 ), type = "Ljung-Box", fitdf = (0+3))


## arima evaluation on orders
orders.resid = resid(sales.final.diff.one)
plot(orders.resid, main = "Residual plot of New orders ( p =  5 & q = 6)")
acf(orders.resid, lag.max = 200,  main = "Residual ACF of New orders ( p =  5 & q = 6)")
pacf(orders.resid, lag.max = 200, main = "Residual PACF of New orders ( p =  5 & q = 6)")
Box.test(orders.resid, lag = ( 5+6+1 ), type = "Ljung-Box", fitdf = (5+6))
```

```{r}
## checking garch order on Total sales with p = 3, q = 0, d= 1
final.bic = Inf
final.order = c(0,0)
sales.test.garch = function(m, n) {
  spec = ugarchspec(variance.model = list(garchOrder = c(m, n)), mean.model = list(armaOrder=c(3,0),include.mean = T), distribution.model = "std")
  fit = ugarchfit(spec, sales.train, solver= "hybrid")
  current.bic = infocriteria(fit)[2]
  bicdata = data.frame(m,n,current.bic)
  names(bicdata) = c("m","n","BIC")
  print(paste(m,n,current.bic, sep=" "))
  return(bicdata)
}

orders = data.frame(Inf, Inf, Inf)
names(orders) = c("m", "n", "BIC")

for (m in 0:2) for (n in 0:2) {
  Error = tryCatch(orders = rbind(orders, sales.test.garch(m, n)), error = function(e) e) 
  if (inherits(Error, "error"))
    next
}
```

```{r}
## checking garch order on New orders  with p = 3, q = 0, d= 1
final.bic = Inf
final.order = c(0,0)
orders.test.garch = function(m, n) {
  spec = ugarchspec(variance.model = list(garchOrder = c(m, n)), mean.model = list(armaOrder=c(5,6),include.mean = T), distribution.model = "std")
  fit = ugarchfit(spec, orders.train, solver= "hybrid")
  current.bic = infocriteria(fit)[2]
  #garchforecast = ugarchforecast(fitORspec = fit, n.ahead = orders.test)
  bicdata = data.frame(m,n,current.bic)
  names(bicdata) = c("m","n","BIC")
  print(paste(m,n,current.bic, sep=" "))
  return(bicdata)
}

order = data.frame(Inf, Inf, Inf)
names(order) = c("m", "n", "BIC")

for (m in 0:2) for (n in 0:2) {
  Error = tryCatch(order = rbind(orders, orders.test.garch(m, n)), error = function(e) e) 
  if (inherits(Error, "error"))
    next
}
```

```{r}
## fit optimal orders in arma garch model (Total sales)
garchspec.sales = ugarchspec(variance.model = list(garchOrder = c(1,0)), mean.model = list(armaOrder = c(3,0),include.mean = T), distribution.model = "std")

garchfit.sales = ugarchfit(garchspec.sales, sales.train, solver="hybrid")

garchforecast.sales = ugarchforecast(fitORspec = garchfit.sales, n.ahead = 100)

## predicted mean
garchmean.sales = fitted(garchfit.sales)
#garchmean

## predicted volatilities
garchvola.sales = sigma(garchfit.sales)
#garchvola
plot(garchvola.sales, main = "volatilities of Total sales")
```

```{r}
## fit optimal orders in arma garch model (New orders)
garchspec.orders = ugarchspec(variance.model = list(garchOrder = c(1,1)), mean.model = list(armaOrder = c(5,6),include.mean = T), distribution.model = "std")

garchfit.orders = ugarchfit(garchspec.orders, orders.train, solver="hybrid")

garchforecast.orders = ugarchforecast(fitORspec = garchfit.orders, n.ahead = 100)

## predicted mean
garchmean.orders = fitted(garchfit.orders)
#garchmean

## predicted volatilities
garchvola.orders = sigma(garchfit.orders)
#garchvola
plot(garchvola.orders, main = "Volatilities of New Orders")
```

```{r}
cor(garchvola.sales,garchvola.orders)
```


### The part below is multivariate time series. 

```{r}
library(data.table)
library(vars)
library(xts)
library(mgcv)
library(stats)
library(tseries)
library(aod)
```


####### 
(1)I checked plots of total sales and new orders from original data. 
(3)Identify max and min values in total sales and new orders. the min values in total sales and new orders happened to be on the same date 2020-04-01.
(2)I gave dataframe a new name data2 (just incase in later we might need some variables that i built under original dataframe called data.), I have removed max and min values in the total sales, the date is 0001-10-01 and 2020-04-01, also removed min value of new orders but not max value of new orders, since the min value appeared during covid, but could not find a reason to remove max value.
(4)Compare the output from the code with the value in the plots of total sales and new orders. they seemed match each other. so i kept the data2 with the extreme values in total sales and new orders removed. 



```{r}
sales.plot = autoplot(sales) + xlab("date") + ylab("sales") + ggtitle("Total Sales")
sales.plot

orders.plot = autoplot(orders) + xlab("data") + ylab("orders") + ggtitle("New Orders")
orders.plot

newplots = cbind(sales, orders)
new.plots = autoplot(newplots, facets = F) + scale_y_continuous(trans = "log10")
new.plots
```


```{r}
##### this chunk of the code below is used to identify and remove extreme points

# here i checked extreme points in the original data without convert them to time series. 
data2 = read.csv("VehicleData-1.csv", header = T)
data2$DATE = as.Date(data$DATE, format =  "%m/%d/%Y")

data2 %>% mutate_all(~replace(., is.nan(.), 0))

tail(data2)
```


```{r}
## identify extreme points in the total sales.
max.sales = data2[which.max(data2$Total.Sales), ]
max.sales

min.sales = data2[which.min(data2$Total.Sales), ]
min.sales
```


```{r}
## identify extreme points in the new orders
max.orders = data2[which.max(data2$New.Orders), ]
max.orders

min.orders = data2[which.min(data2$New.Orders), ]
min.orders
```


```{r}
## remove those identified extreme points from total sales and new orders

## first to remove one max and two min values of tatal sales from dataframe, which is (date) 0001-10-01 and 2020-04-01
data2 = data2[data2$Total.Sales < 22.055,	]
data2 = data2[data2$Total.Sales > 11.736, ]

## first to remove two min values of new orders from dataframe, which is (date) 2020-04-01 since it was during Covid. 

data2 = data2[data2$New.Orders < 64455, ]

data2 = data2[data2$New.Orders > 24561	,	]

data2

str(data2)
```

```{r}
## now convert the data with extreme values removed to time series. 
new.sales = ts(data2$Total.Sales, start = c(2000,1,1), frequency = 12)
new.orders = ts(data2$New.Orders, start = c(2000,1,1), frequency = 12)

```

```{r}
new.sales.plot = autoplot(new.sales) + xlab("date") + ylab("sales") + ggtitle("Total Sales") 
new.sales.plot

new.orders.plot = autoplot(new.orders) + xlab("data") + ylab("orders") + ggtitle("New Orders")
new.orders.plot

```

































































































































































































































